{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-08-14 thesis/tm.tex
nother tm result
more txn bugs
challenge unconscious bias

If these instructions from the {\tt x++} in the transaction are preempted,
with another thread's access to {\tt x} interleaved in between,
the transaction will abort.
So, the interleaving
\[
	\tiat{1}, \tiat{2a}, \tiat{2b}, \tjat{1}, \tjat{2}, \tjat{3}, \tiat{2c}, \tiat{3}
\]
or, henceforth abbreviated for clarity:
\[
	\tiat{1-2b}, \tjat{1-3}, \tiat{2c-3}
\]
is not possible; rather, \ti will fall into the backup path:
\[
	\tiat{1-2b}, \tjat{1-3}, \tiat{4-7}
\]
However, the {\tt x++} operation from the failure path (correspondingly $6a$, $6b$, $6c$)
{\em can} be thusly separated with conflicting accesses interleaved in between,
since the mutex only protects the failure path against other failure paths,
but not against the transaction itself.
So (assuming {\tt x} is intended to be a precise counter rather than a sloppy one),
%losing one of the increments to which constitutes a bug),
the following interleaving exposes a bug:%
\footnote{Note also that this bug requires either at least 3 threads or at least 2 iterations between 2 threads to expose;
this highlights MC's dependence on its test cases to produce meaningful state spaces in the first place.}
\[
	\tiat{1-2b}, \tjat{1-3}, \tiat{4-6b}, \tkat{1-3}, \tiat{6c-7}
\]
Prior work \cite{tm-benchmark-cmu} proposed the idiom shown in Figure~\ref{fig:htm-fixed}
to exclude this family of interleavings,
which shows that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
%further motivating our research.

@@

\begin{itemize}
	\item $\tiiat{\alpha}$ be an HTM begin operation,
	\item $\tiiat{\beta_1}\dots\tiiat{\beta_n}$ be the transaction body (with $\beta_n$ the HTM end call),
	\item $\tiiat{\phi_1}\dots\tiiat{\phi_m}$ be the failure path, and
	\item $\tiiat{\omega_1}\dots\tiiat{\omega_l}$ be the subsequent code executed unconditionally.
\end{itemize}
Note that arbitrary code may not be structured to distinguish these as nicely as in our examples;
e.g., more code may exist in the success branch after {\tt \_xend()};
such would be considered part of $\omega$ here.

@@

Up to four abort codes are considered depending on the testing options listed in the previous section.
All abort codes are simultaneously added to both lists,
unless already present on {\tt xabort\_codes\_ever},
in which case not added to {\tt xabort\_codes\_todo} to avoid duplicate work.
\begin{itemize}
	\item {\tt \_XABORT\_RETRY}:
		When a {\tt xbegin} preemption point is created ({\tt save\_setjmp()}),
		if {\tt -S} is not set,
		both {\tt xabort\_codes} lists are initialized to contain this code.
		This represents the possibility for a hardware transaction to fail
		for reasons outside of the programmer's control such as system interrupts or cache eviction.
		If {\tt -S} is set they are initialized to empty,
		representing either STM's policy of failing only when a true memory conflict arises,
		or a TSX user wrapping all their {\tt \_xbegin()}s in a retry loop such as in \cite{sigbovik-htm}.
	\item {\tt \_XABORT\_CONFLICT}:
		Whenever DPOR detects a memory conflict between two transitions ({\tt shimsham\_shm()}),
		if {\tt -A} is set,
		% for each of the two which are within a transaction, % no!!!!!!!
		%it finds the nearest ancestor {\tt xbegin} preemption point by the same thread
		if the later-executed transition is a transactional success path,
		it adds this code to both lists of the immediately preceding preemption point.
		Note that a transaction should suffer a conflict abort
		only when executed {\em after} a conflicting memory access
		to avoid circular causality
		(see {\tt 410user/progs/htm\_causality.c} for rationale).
		Should the transaction happen to be executed first,
		DPOR will first try to reorder it as normal,
		and then abort it as described if the conflict persists.
	\item {\tt \_XABORT\_EXPLICIT}:
		Whenever the program invokes {\tt \_xabort()} ({\tt sched\_update\_user\_\allowbreak{}state\_machine()}),
		if {\tt -A} is set,
		this code is added to both lists, bitwise-ored with the user-supplied argument code
		as specified in \cite{htm-gcc},
		and execution of the transactional path immediately stops.
	\item {\tt \_XABORT\_CAPACITY}:
		Whenever the program invokes a system call during a transaction,
		if {\tt -A} is set,
		this code is added to both lists,
		and execution of the transactional path immediately stops.
\end{itemize}

@@

			& 2,1	&\cpu{104.37}&\wtm{17.00}&\ints{*6}& \cpu{*43.76}& \wtm{*12.51}& \ints{33	}& \ETA{73} \\

@@

			& 2,2	&\cpu{ 66.47}&\wtm{*12.42}&\ints{*38}& \cpu{*45.22}& \wtm{15.30}& \ints{85	}& \ETA{860} \\
			& 2,3	&\cpu{ 99.41}&\wtm{*15.92}&\ints{*48}& \cpu{*53.57}& \wtm{21.34}& \ints{217	}& \ETA{9120} \\
			& 2,4	&\cpu{125.23}&\wtm{*19.46}&\ints{*68}& \cpu{*71.01}& \wtm{38.86}& \ints{537	}& \ETA{91239} \\

@@

			& 3,1	&\cpu{313.36}&\wtm{*44.50}&\ints{*34}&\cpu{*78.23}& \wtm{60.49}& \ints{1016	}& \ETA{3543} \\
			& 3,2	&\cpu{*6500.36}&\wtm{*816.27}&\ints{*262}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{2161715} \\

@@

Regarding the spinlock bug,
as HTM is fundamentally a performance-minded concurrency extension,
users are likely to care about more probabilistic properties of their code,
such as requiring a transaction abort rate below a certain threshhold
owing to the nature of its workload.
Landslide cannot in general test for performance degradation bugs,
because all interleavings are equal in Landslide's eyes, and probability is no object.
However, {\tt lock\_fast} illustrates that model checking can still check some interesting performance properties
as long as the element of probability can be removed.
Future work may attempt to verify a wider range of performance properties,
but with a hybrid approach between model checking and what other technique is not yet known.

@@

Table~\ref{tab:tm-verifs2} shows the new resulting levels of verification
Landslide reached before the same 10-hour timeout.
Provided that one trusts the {\tt lock} tests correctly check the desired properties,
and that open-coding hadn't introduced any new bugs (such as Figure~\ref{fig:avlbug}'s),
the benefit is clear:
the data structure tests' state spaces become much more tractable,
%the data structure tests suddenly reach unprecedented thread/iteration counts with ease,
% 'so on' here includes free/re-malloc mem conflicts; see comment in avl_mutex.cpp
their state space growth now defined only by their internal conflicts from tree rebalancing, map resizing, and so on.
In total,
summing the testing times of {\tt lock}({\tt mutex})$(K,N)$ and {\tt avl\_mutex}$(K,N)$
produces the same verification as {\tt avl\_fixed}$(K,N)$ far more cheaply.
Furthermore, {\tt lock}'s verification can be reused,
whereas {\tt avl\_fixed} and {\tt map\_basic} effectively duplicated the mutex verification between them.

\begin{table}[t]
	\begin{center}
		\footnotesize
		\begin{tabular}{cc}
		\begin{tabular}{cc||r|r}
			& & \multicolumn{2}{c}{\bf STM ({\tt -A -S})} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf params & \em (or \ETAdag{\bf \em est.}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.41} & \ints{4} \\
			({\tt spinlock})
			& 2,2 & \cpu{198.77} & \ints{1702} \\
			% if lock(spin_fixed) will time out on 2,3 and 3,2, this is guaranteed to aswell
			& 3,1 & \cpu{35.28} & \ints{246} \\
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.55} & \ints{4} \\
			({\tt spin\_fixed})
			& 2,2 & \cpu{105.57} & \ints{998} \\
			& 2,3 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 3,1 & \cpu{28.27} & \ints{186} \\
			& 3,2 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 4,1 & \ETAdag{16h 26m} & \ETAdag{432628} \\ % was: 30408, 282084 pre mkrun
			% [JOB 4] progress: 286910/432628 brs (66.317896%), ETA 6h 26m 25s (elapsed 9h 59m 48s)
			% total CPU time consumed: 10h 4s (36004288982 usecs) (core saturation: 12%)
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.44} & \ints{4}		\\
			({\tt mutex})
			& 2,2 & \cpu{16.83} & \ints{180}	\\
			& 2,3 & \cpu{405.21} & \ints{9372}	\\
			& 2,4 & \cpu{qqq} & \ints{qqq}	\\ % was: 15266 sec 384512 pre mkrun
			& 3,1 & \cpu{15.00} & \ints{132}	\\
			%& 3,2 & \cpu{29432.87} & \ints{707086}	\\ % pre mkrun fix (let alone bottom1)
			& 3,2 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 4,1 & \cpu{665.89} & \ints{15064}	\\
			\hline
			{\tt lock\_fast}
			& 2,1 & \cpu{3.25} & \ints{1}	\\
			({\tt spin\_fixed})
			& 9,9 & \cpu{4.61} & \ints{1}	\\
			\hline
			{\tt lock\_fast}
			& 2,1 & \cpu{3.19} & \ints{1}	\\
			({\tt mutex})
			& 9,9 & \cpu{4.62} & \ints{1}	\\
			\end{tabular}
			&
			\begin{tabular}{cc||r|r}
			& & \multicolumn{2}{c}{\bf Non-transactional} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf params & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt avl\_mutex}
			& 2,1 & \cpu{3.48} & \ints{7} \\
			& 2,2 & \cpu{6.25} & \ints{85} \\
			& 2,3 & \cpu{24.69} & \ints{561} \\
			& 2,4 & \cpu{217.98} & \ints{4984} \\
			& 3,1 & \cpu{8.26} & \ints{129} \\
			& 3,2 & \cpu{1403.46} & \ints{30653} \\
			& 3,3 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 4,1 & \cpu{199.96} & \ints{4488} \\
			& 4,2 & \ETAdag{qqq} & \ETAdag{qqq} \\
			\cline{1-4}
			{\tt map\_mutex}
			& 2,1 & \cpu{39.81} & \ints{83} \\
			& 2,2 & \ETAdag{26h 21m} & \ETAdag{1085126} \\
			% [JOB 9] progress: 496530/1085126 brs (45.757812%), ETA 16h 21m 0s (elapsed 9h 59m 38s)
			% total CPU time consumed: 10h 19s (36019507965 usecs) (core saturation: 12%)
			& 3,1 & \ETAdag{173d 17h} & \ETAdag{12572187} \\
			% [JOB 9] progress: 496281/12572187 brs (3.947451%), ETA 173d 6h 38m 55s (elapsed 9h 59m 43s)
			% total CPU time consumed: 10h 20s (36020148608 usecs) (core saturation: 12%)
			\\
			\\
			\\
			\\
			\\
		\end{tabular}
		\\
		\\
		(a) Verifying HTM locks alone.
		&
		(b) Verifying the lock's client code.
		\end{tabular}
	\end{center}
	\caption{Continuation of Table~\ref{tab:tm-verifs},
		demonstrating abstraction reduction \cite{dbug-phdthesis}
		on the {\tt avl\_fixed} and {\tt map\_basic} tests
		by verifying HTM mutex implementations separately.
		Tested with STM semantics, as both lock implementations include a retry loop.}
	\label{tab:tm-verifs2}
\end{table}

@@

% "Although we skipped implementing it to begin with, the sheer size of this reduction also
% reflects on the numbers one might see betwee the original reduction in theorem so-and-so that lets us
% treat transactional success paths as mutually-exclusive (and, once started, abort-proof) to begin with"
% this sentence is a complete mess....

@@

On the other hand, some programs may clearly annotate their intention
for abort paths to be executed only in case of actual memory conflicts.
{\tt swap}, {\tt avl\_insert}, and {\tt map\_basic} abstract their {\tt \_xbegin()} calls
behind an interface which can be implemented either with or without retry loops,
while {\tt lock}({\tt spinlock}) and {\tt lock}({\tt mutex}) implement the retry loop directly.
In these cases, the user can assure themself of STM reduction's soundness by visual inspection.
In fact, Landslide's current implementation gets stuck in infinitely deep interleavings
whenever it encounters a retry loop
% not actually sure why, as it should be around-a-pp and not on the 1st branch...
(bypassing even its heuristic infinite loop detection),
so for now the user must inspect the test case to determine which testing mode to use.
Future work could automatically identify a program's retry loops
and give up on {\tt \_XABORT\_RETRY} by switching to STM mode on-the-fly,
much like Landslide's heuristic synchronization detection does for yield loops (\sect{\ref{sec:landslide-blocking}}).

>>>>

If these instructions from the {\tt x++} in the transaction are preempted,
with another thread's access to {\tt x} interleaved in between,
the transaction will abort.
So, the interleaving
\[
	\tiat{1}, \tiat{2a}, \tiat{2b}, \tjat{1}, \tjat{2}, \tjat{3}, \tiat{2c}, \tiat{3}
\]
or, henceforth abbreviated for clarity:
\[
	\tiat{1-2b}, \tjat{1-3}, \tiat{2c-3}
\]
is not possible; rather, \ti will fall into the backup path:
\[
	\tiat{1-2b}, \tjat{1-3}, \tiat{4-7}
\]
However, the {\tt x++} operation from the failure path (correspondingly $6a$, $6b$, $6c$)
{\em can} be thusly separated with conflicting accesses interleaved in between,
since the mutex only protects the failure path against other failure paths,
but not against the transaction itself.
So (assuming {\tt x} is intended to be a precise counter rather than a sloppy one),
%losing one of the increments to which constitutes a bug),
the following interleaving exposes a bug:%
\footnote{Note also that this bug requires either at least 3 threads or at least 2 iterations between 2 threads to expose;
this highlights MC's dependence on its test cases to produce meaningful state spaces in the first place.}
\[
	\tiat{1-2b}, \tjat{1-3}, \tiat{4-6b}, \tkat{1-3}, \tiat{6c-7}
\]
Prior work \cite{tm-benchmark-cmu} proposed the idiom shown in Figure~\ref{fig:htm-fixed}
to exclude this family of interleavings,
which shows that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
%further motivating the research.

@@

\begin{itemize}
	\item $\tiiat{\alpha}$ be an HTM begin operation,
	\item $\tiiat{\beta_1}\dots\tiiat{\beta_n}$ be the transaction body (with $\beta_n$ the HTM end call),
	\item $\tiiat{\phi_1}\dots\tiiat{\phi_m}$ be the failure path, and
	\item $\tiiat{\omega_1}\dots\tiiat{\omega_l}$ be the subsequent code executed unconditionally.
\end{itemize}
Note that arbitrary code may not be structured to distinguish these as nicely as in the examples;
e.g., more code may exist in the success branch after {\tt \_xend()};
such would be considered part of $\omega$ here.

@@

Up to four abort codes are considered depending on the testing options listed in the previous section.
All abort codes are simultaneously added to both lists,
unless already present on {\tt xabort\_codes\_ever},
in which case not added to {\tt xabort\_codes\_todo} to avoid duplicate work.
\begin{itemize}
	\item {\tt \_XABORT\_RETRY}:
		When a {\tt xbegin} preemption point is created ({\tt save\_setjmp()}),
		if {\tt -S} is not set,
		both {\tt xabort\_codes} lists are initialized to contain this code.
		This represents the possibility for a hardware transaction to fail
		for reasons outside of the programmer's control such as system interrupts or cache eviction.
		If {\tt -S} is set they are initialized to empty,
		representing either STM's policy of failing only when a true memory conflict arises,
		or a TSX user wrapping all her {\tt \_xbegin()}s in a retry loop such as in \cite{sigbovik-htm}.
	\item {\tt \_XABORT\_CONFLICT}:
		Whenever DPOR detects a memory conflict between two transitions ({\tt shimsham\_shm()}),
		if {\tt -A} is set,
		% for each of the two which are within a transaction, % no!!!!!!!
		%it finds the nearest ancestor {\tt xbegin} preemption point by the same thread
		if the later-executed transition is a transactional success path,
		it adds this code to both lists of the immediately preceding preemption point.
		Note that a transaction should suffer a conflict abort
		only when executed {\em after} a conflicting memory access
		to avoid circular causality
		(see {\tt 410user/progs/htm\_causality.c} for rationale).
		Should the transaction happen to be executed first,
		DPOR will first try to reorder it as normal,
		and then abort it as described if the conflict persists.
	\item {\tt \_XABORT\_EXPLICIT}:
		Whenever the program invokes {\tt \_xabort()} ({\tt sched\_update\_user\_\allowbreak{}state\_machine()}),
		if {\tt -A} is set,
		this code is added to both lists, bitwise-ored with the user-supplied argument code
		as specified in \cite{htm-gcc},
		and execution of the transactional path immediately stops.
	\item {\tt \_XABORT\_CAPACITY}:
		Whenever the program invokes a system call during a transaction,
		if {\tt -A} is set,
		this code is added to both lists,
		and execution of the transactional path immediately stops.
\end{itemize}

@@

			& 2,1	&\cpu{104.37}&\wtm{17.00}&\ints{*6}& \cpu{*43.76}& \wtm{*12.51}& \ints{33}	& \ETA{73} \\
			% TODO: for maximal:
			% 33 interleavings tested; 1 preemptions; job time 4s; pldi time 12326606; new-fixed pldi cputime 43294461
			% 73 brs eta

@@

			& 2,2	&\cpu{ 66.47}&\wtm{*12.42}&\ints{*38}& \cpu{*45.22}& \wtm{15.30}& \ints{85}	& \ETA{860} \\
			& 2,3	&\cpu{ 99.41}&\wtm{*15.92}&\ints{*48}& \cpu{*53.57}& \wtm{21.34}& \ints{217}	& \ETA{9120} \\
			& 2,4	&\cpu{*38.92}&\wtm{*8.61}&\ints{*34}& \cpu{82.95}& \wtm{48.31}& \ints{537}	& \ETA{91239} \\
			% 34 interleavings tested; 1 preemptions; job time 6s; pldi time 8614371; new-fixed pldi cputime 38917975
			% 537 interleavings tested; 1 preemptions; job time 38s; pldi time 48314592; new-fixed pldi cputime 82952352

@@

			& 3,1	&\cpu{*38.59}&\wtm{*8.50}&\ints{*32}&\cpu{88.28}& \wtm{72.41}& \ints{1016}	& \ETA{3543} \\
			% 32 interleavings; 258 preemptions; job time 5s; pldi time 8506344; new-fixed pldi cputime 38594745
			% 1016 interleavings; 258 preemptions; job time 58s; pldi time 72406995; new-fixed pldi cputime 88281009
			& 3,2	&\cpu{*1572.83}&\wtm{*199.98}&\ints{*262}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{1683509} \\
			% 262 interleavings; 260 preemptions; job time 47s; pldi time 199982231; new-fixed pldi cputime 1572825801
			% [JOB 8] progress: 47047/1683509 brs (2.794578%), ETA 1y 237d 20h 41m 57s (elapsed 59m 44s)

@@

Regarding the spinlock bug,
as HTM is fundamentally a performance-minded concurrency extension,
the user may also care about more probabilistic properties of her code,
such as requiring a transaction abort rate below a certain threshhold
owing to the nature of its workload.
Landslide cannot in general test for performance degradation bugs,
because all interleavings are equal in Landslide's eyes, and probability is no object.
However, {\tt lock\_fast} illustrates that model checking can still check some interesting performance properties
as long as the element of probability can be removed.
Future work may attempt to verify a wider range of performance properties,
but with a hybrid approach between model checking and what other technique is not yet known.

@@

Table~\ref{tab:tm-verifs2} shows the new resulting levels of verification
Landslide reached before the same 10-hour timeout.
Provided that one trusts the {\tt lock} tests correctly check the desired properties,
and that open-coding hadn't introduced any new bugs (such as Figure~\ref{fig:avlbug}'s),
the benefit is clear:
the data structure tests' state spaces become much more tractable,
%the data structure tests suddenly reach unprecedented thread/iteration counts with ease,
% 'so on' here includes free/re-malloc mem conflicts; see comment in avl_mutex.cpp
their state space growth now defined only by internal conflicts from tree rebalancing, map resizing, and so on.
In total,
summing the testing times of {\tt lock}({\tt mutex})$(K,N)$ and {\tt avl\_mutex}$(K,N)$
produces the same verification as {\tt avl\_fixed}$(K,N)$ far more cheaply.
Furthermore, {\tt lock}'s verification can be reused,
whereas {\tt avl\_fixed} and {\tt map\_basic} effectively duplicated the mutex verification between them.

\begin{table}[t]
	\begin{center}
		\footnotesize
		\begin{tabular}{cc}
		\begin{tabular}{cc||r|r}
			& & \multicolumn{2}{c}{\bf STM ({\tt -A -S})} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf params & \em (or \ETAdag{\bf \em est.}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.41} & \ints{4} \\
			({\tt spinlock})
			& 2,2 & \cpu{198.77} & \ints{1702} \\
			% if lock(spin_fixed) will time out on 2,3 and 3,2, this is guaranteed to aswell
			& 3,1 & \cpu{35.28} & \ints{246} \\
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.55} & \ints{4} \\
			({\tt spin\_fixed})
			& 2,2 & \cpu{105.57} & \ints{998} \\
			& 2,3 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 3,1 & \cpu{28.27} & \ints{186} \\
			& 3,2 & \ETAdag{13y 281d} & \ETAdag{1443676} \\
			% [JOB 4] progress: 169893/1443676 brs (11.768080%), ETA 13y 280d 17h 30m 54s (elapsed 9h 59m 48s)
			& 4,1 & \ETAdag{16h 26m} & \ETAdag{432628} \\ % was: 30408, 282084 pre mkrun
			% [JOB 4] progress: 286910/432628 brs (66.317896%), ETA 6h 26m 25s (elapsed 9h 59m 48s)
			% total CPU time consumed: 10h 4s (36004288982 usecs) (core saturation: 12%)
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.44} & \ints{4}		\\
			({\tt mutex})
			& 2,2 & \cpu{16.83} & \ints{180}	\\
			& 2,3 & \cpu{405.21} & \ints{9372}	\\
			& 2,4 & \cpu{qqq} & \ints{qqq}	\\ % was: 15266 sec 384512 pre mkrun
			& 3,1 & \cpu{15.00} & \ints{132}	\\
			%& 3,2 & \cpu{29432.87} & \ints{707086}	\\ % pre mkrun fix (let alone bottom1)
			& 3,2 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 4,1 & \cpu{665.89} & \ints{15064}	\\
			\hline
			{\tt lock\_fast}
			& 2,1 & \cpu{3.25} & \ints{1}	\\
			({\tt spin\_fixed})
			& 9,9 & \cpu{4.61} & \ints{1}	\\
			\hline
			{\tt lock\_fast}
			& 2,1 & \cpu{3.19} & \ints{1}	\\
			({\tt mutex})
			& 9,9 & \cpu{4.62} & \ints{1}	\\
			\end{tabular}
			&
			\begin{tabular}{cc||r|r}
			& & \multicolumn{2}{c}{\bf Non-transactional} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf params & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt avl\_mutex}
			& 2,1 & \cpu{3.48} & \ints{7} \\
			& 2,2 & \cpu{6.25} & \ints{85} \\
			& 2,3 & \cpu{24.69} & \ints{561} \\
			& 2,4 & \cpu{217.98} & \ints{4984} \\
			& 3,1 & \cpu{8.26} & \ints{129} \\
			& 3,2 & \cpu{1403.46} & \ints{30653} \\
			& 3,3 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 4,1 & \cpu{199.96} & \ints{4488} \\
			& 4,2 & \ETAdag{qqq} & \ETAdag{qqq} \\
			\cline{1-4}
			{\tt map\_mutex}
			& 2,1 & \cpu{39.81} & \ints{83} \\
			& 2,2 & \ETAdag{26h 21m} & \ETAdag{1085126} \\
			% [JOB 9] progress: 496530/1085126 brs (45.757812%), ETA 16h 21m 0s (elapsed 9h 59m 38s)
			% total CPU time consumed: 10h 19s (36019507965 usecs) (core saturation: 12%)
			& 3,1 & \ETAdag{173d 17h} & \ETAdag{12572187} \\
			% [JOB 9] progress: 496281/12572187 brs (3.947451%), ETA 173d 6h 38m 55s (elapsed 9h 59m 43s)
			% total CPU time consumed: 10h 20s (36020148608 usecs) (core saturation: 12%)
			\\
			\\
			\\
			\\
			\\
		\end{tabular}
		\\
		\\
		(a) Verifying HTM locks alone.
		&
		(b) Verifying the lock's client code.
		\end{tabular}
	\end{center}
	\caption{Continuation of Table~\ref{tab:tm-verifs},
		demonstrating abstraction reduction \cite{dbug-phdthesis}
		on the {\tt avl\_fixed} and {\tt map\_basic} tests
		by verifying HTM mutex implementations separately.
		Tested with STM semantics, as both lock implementations include a retry loop.}
	\label{tab:tm-verifs2}
\end{table}

@@

% "Although I skipped implementing it to begin with, the sheer size of this reduction also
% reflects on the numbers one might see betwee the original reduction in theorem so-and-so that lets us
% treat transactional success paths as mutually-exclusive (and, once started, abort-proof) to begin with"
% this sentence is a complete mess....

@@

On the other hand, some programs may clearly annotate their intention
for abort paths to be executed only in case of actual memory conflicts.
{\tt swap}, {\tt avl\_insert}, and {\tt map\_basic} abstract their {\tt \_xbegin()} calls
behind an interface which can be implemented either with or without retry loops,
while {\tt lock}({\tt spinlock}) and {\tt lock}({\tt mutex}) implement the retry loop directly.
In these cases, the user can assure herself of STM reduction's soundness by visual inspection.
In fact, Landslide's current implementation gets stuck in infinitely deep interleavings
whenever it encounters a retry loop
% not actually sure why, as it should be around-a-pp and not on the 1st branch...
(bypassing even its heuristic infinite loop detection),
so for now the user must inspect the test case to determine which testing mode to use.
Future work could automatically identify a program's retry loops
and give up on {\tt \_XABORT\_RETRY} by switching to STM mode on-the-fly,
much like Landslide's heuristic synchronization detection does for yield loops (\sect{\ref{sec:landslide-blocking}}).
